Capstone Movielens Report

### 2025-11-21 reports/capstone-movielens.site/05.appendix.Rmd
Update  for function: `model.tune.param_range` (Detais section, steps 7,8).

#### [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#LX) Function {#func.get_fine_tune.param.endpoints.idx}
\

@@

```{r eval=FALSE}
  preset.result.idx <- get_fine_tune.param.endpoints.idx(preset.result)

@@

::: { #func.get_fine_tune.param.endpoints.idx.args. .sidebar }
  - **preset.result:** ;

:::
::: { #func.get_fine_tune.param.endpoints.idx.args. .sidebar }
  - **:** ;

::: { #func.get_fine_tune.param.endpoints.idx.details .sidebar }
##### Details
\

::: { #func.get_fine_tune.param.endpoints.idx.value .sidebar }
##### Value
\

@@

The source code of the [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#LX) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#LX) function is also available in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L69) section of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#LX) script on _GitHub_.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above, we use the custom function [put_log1](#func.put_log2) described above in the [Logging Functions] section of *This Appendix*.
:::
:::

@@

::: { #func.model.tune.param_range.details.step6 .sidebar }
  6. Calls the helper function [tune.model_param](#func.tune.model_param), passing the vector created earlier in [step 3](#func.model.tune.param_range.details.step3) to the [param_values](#func.tune.model_param.args.param_values) argument (along with the values of the rest arguments: [fn_tune.test.param_value](#func.tune.model_param.args.fn_tune.test.param_value), [break.if_min](#func.tune.model_param.args.break.if_min), [steps.beyond_min](#func.tune.model_param.args.steps.beyond_min)) of the called function and stores the returned result in the local variable `tuned_result`:
```{r eval=FALSE}
      tuned_result <- tune.model_param(test_param_vals, 
                                        fn_tune.test.param_value,
                                        break.if_min,
                                        steps.beyond_min)
```

::: { #func.model.tune.param_range.details.step7 .sidebar }
  7. Extracts the internal data frame stored in the variable `tuned_result$tuned.result` from the data returned by the [tune.model_param](#func.tune.model_param) helper function (and saved in the local variable `tuned_result`) in the [previous step](#func.model.tune.param_range.details.step6) and stores it in the local variable `tuned.result`. Then stores `tuned_result` along with the necessary auxiliary information in the cache file with the full name created earlier in [step 4](#func.model.tune.param_range.details.step4):
```{r eval=FALSE}
      tuned.result <- tuned_result$tuned.result

@@

::: { #func.model.tune.param_range.details.step8 .sidebar }
Next, the function figures out the boundary indices of a range of values from the neighborhood of the minimum `RMSE` by calling internally another helper function [get_fine_tune.param.endpoints.idx](#func.get_fine_tune.param.endpoints.idx) (see the description below) and, using one more helper function [get_best_param.result](#func.get_best_param.result)  (see the description below), a pair of values,  corresponding to the best result: minimal  `RMSE` and corresponding input parameter value (which is considered the best).

@@

@@

::: {#func.model.tune.param_range.code .sidebar }
##### Source Code
\
Below is the simplified version of the source code of the [model.tune.param_range](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L198) function:
```{r eval=FALSE}
model.tune.param_range <- function(loop_starter,
                             tune_dir_path,
                             cache_file_base_name,
                             fn_tune.test.param_value,
                             # interval_divisor.multiplier = 4,
                             max.identical.min_RMSE.count = 4,
                             endpoint.min_diff = 0, #1e-07,
                             break.if_min = TRUE,
                             steps.beyond_min = 2){

@@

#### [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L87) Function {#func.get_fine_tune.param.endpoints.idx}
\

##### Signature
\
```{r eval=FALSE}
get_fine_tune.param.endpoints.idx <- function(preset.result) {
  # ...

  c(start = i, 
    end = j,
    best = best_RMSE.idx)
}
```

::: {#func.get_fine_tune.param.endpoints.idx.args.preset.result .sidebar }
###### ***preset.result***
\
A data structure compatible with the `tuned.result` item of the data structure returned by the [tune.model_param](#func.tune.model_param) function (see the [Value](#func.tune.model_param.value) subsection of the `tune.model_param` function description section for more details). Here is a sample of using this parameter in the `model.tune.param_range` function [code](#func.model.tune.param_range.code):
```{r eval=FALSE}
    tuned_result <- tune.model_param(test_param_vals, 
                                  fn_tune.test.param_value,
                                  break.if_min,
                                  steps.beyond_min)

    tuned.result <- tuned_result$tuned.result
    # tuned.result = data.frame(RMSE, parameter.value)
    # ...
    bound.idx <- get_fine_tune.param.endpoints.idx(tuned.result)
    # bounnd.idx =  c(start.index, end.index, min_RMSE.index)
```
:::

##### Details
\
The algorithm finds the indices of the nearest neighboring elements of the element corresponding to the minimum RMSE value in the given sequence.

##### Value
\
A vector containing the index of the element corresponding to the minimum `RMSE` value in a given sequence, along with the indices of its nearest neighboring elements:
```{r eval=FALSE}
  c(start, # interval start index
    end,   # interval end index
    best)  # index of the best `RMSE`
```

##### Source Code
\
The source code of the [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L87) function is shown below:
```{r eval=FALSE}
get_fine_tune.param.endpoints.idx <- function(preset.result) {
  best_RMSE <- min(preset.result$RMSE)
  best_RMSE.idx <- which.min(preset.result$RMSE)
  # best_lambda <- preset.result$parameter.value[best_RMSE.idx]

>>>>

#### [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L81) Function {#func.get_fine_tune.param.endpoints.idx}
\

Implements an algorithm that finds the pair of indices of the nearest neighboring elements with values greater than the minimum to the left and right of the element (or group of elements) with the minimum value in the *RMSE* column of the data frame passed in the [preset.result](#func.get_fine_tune.param.endpoints.idx.args.preset.result) argument.

::: {.noteblock data-latex=""}
The algorithm assumes that the set of values in the *RMSE* column is a monotonically non-increasing function of the index until the minimum is reached and monotonically non-decreasing thereafter. That is, it is assumed that the function can have only one minimum on the given set.
:::

@@

::: {.noteblock data-latex=""}
The element named `tuned.result` of the *list* returned by the [tune.model_param](#func.tune.model_param) function described below is a *data frame* suitable for use as the value of the [preset.result](#func.get_fine_tune.param.endpoints.idx.args.preset.result) argument.
:::

The following code snippet shows how to use the data returned by the [tune.model_param](#func.tune.model_param) function as the value of the [preset.result](#func.get_fine_tune.param.endpoints.idx.args.preset.result) argument:
```{r eval=FALSE}
cv.UME.preset.result <- 
  tune.model_param(lambdas, regularize.test_lambda.UM_effect.cv)

```
```{r}
cv.UME.preset.result$tuned.result |>
  data.plot(title = "UME Model Regularization: `lambda` Range Pre-configuration",
              xname = "parameter.value",
              yname = "RMSE",
              xlabel = "Regularization parameter (`lambda`)",
              ylabel = "RMSE")

preset.result <- cv.UME.preset.result$tuned.result
preset.result.idx <- get_fine_tune.param.endpoints.idx(preset.result)
print(preset.result.idx)
```

@@

::: {#func.get_fine_tune.param.endpoints.idx.args.preset.result .sidebar }
  - **preset.result:** Data frame that has an *RMSE* column containing the `RMSE` values corresponding to the *regularization parameter* $\lambda$ values for the model being regularized, as described in the [Details](#func.tune.model_param.details) section of the [tune.model_param](#func.tune.model_param) function description below.

::: { #func.get_fine_tune.param.endpoints.idx.value .sidebar }
##### Value
\
Vector containing the index of the element corresponding to the minimum `RMSE` value in a given sequence, along with the indices of its nearest neighboring elements:
```{r eval=FALSE}
  c(start, # interval start index
    end,   # interval end index
    best)  # index of the best `RMSE`
```

In the example shown above in the [Usage](#func.get_fine_tune.param.endpoints.idx.usage) section we can see the value return by *this function*:
```{r}
print(preset.result.idx)
```

@@

The source code of the [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L81) function is shown below:
```{r eval=FALSE}
get_fine_tune.param.endpoints.idx <- function(preset.result) {
  best_RMSE <- min(preset.result$RMSE)
  best_RMSE.idx <- which.min(preset.result$RMSE)
  # best_lambda <- preset.result$parameter.value[best_RMSE.idx]

::: {.noteblock data-latex=""}
The source code of the [get_fine_tune.param.endpoints.idx](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L81) function is also available in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L69) section of the [common-helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L81) script on _GitHub_.
:::
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above, we use the custom function [put_log1](#func.put_log1) described above in the [Logging Functions] section of *This Appendix*.
:::
:::

@@

::: { #func.model.tune.param_range.details.step6 .sidebar }
  6. Calls the helper function [tune.model_param](#func.tune.model_param) and stores the returned result in the local variable `tuned_result`, passing the vector created earlier in [step 3](#func.model.tune.param_range.details.step3) to the argument [param_values](#func.tune.model_param.args.param_values) along with the values of the remaining arguments passed in the following arguments of the same name to *this function*:
    - [fn_tune.test.param_value](#func.model.tune.param_range.args.fn_tune.test.param_value);
    - [break.if_min](#func.model.tune.param_range.args.break.if_min);
    - [steps.beyond_min](#func.model.tune.param_range.args.steps.beyond_min):
```{r eval=FALSE}
      tuned_result <- tune.model_param(test_param_vals, 
                                        fn_tune.test.param_value,
                                        break.if_min,
                                        steps.beyond_min)
```

::: { #func.model.tune.param_range.details.step7 .sidebar }
  7. Extracts the data frame element stored in the variable `tuned_result$tuned.result` from the *list object* saved in the local variable `tuned_result` in the [previous step](#func.model.tune.param_range.details.step6) and stores it in the local variable `tuned.result`. Then stores the variable `tuned_result` along with the necessary auxiliary information in the cache file with the full name created earlier in [step 4](#func.model.tune.param_range.details.step4):
```{r eval=FALSE}
      tuned.result <- tuned_result$tuned.result

@@

::: { #func.model.tune.param_range.details.step8 .sidebar }
8. Determines the endpoints for a new interval of $\lambda$ values in the neighborhood of the minimum *RMSE* value figured out so far, with use of the helper function [get_fine_tune.param.endpoints.idx](#func.get_fine_tune.param.endpoints.idx) described above, to improve the accuracy of the *RMSE* calculation in the next iteration, and stores them in the `seq_start` and `seq_end` local variables:
```{r eval=FALSE}
    bound.idx <- get_fine_tune.param.endpoints.idx(tuned.result)
    start.idx <- bound.idx["start"]
    end.idx <- bound.idx["end"]

    seq_start <- prm_val.leftmost.tmp
    seq_end <- prm_val.rightmost.tmp
```
:::

@@

```{r eval=FALSE}
    if (RMSE.leftmost.tmp - min_RMSE >= endpoint.min_diff) {
      prm_val.leftmost <- prm_val.leftmost.tmp
      RMSE.leftmost <- RMSE.leftmost.tmp
      # browser()
    }

    if (RMSE.rightmost.tmp - min_RMSE >= endpoint.min_diff) {
      prm_val.rightmost <- prm_val.rightmost.tmp
      RMSE.rightmost <- RMSE.rightmost.tmp
      # browser()
    } 
```
:::

```{r eval=FALSE}
```
:::
::: { #func.model.tune.param_range.details.step .sidebar }

@@

::: {#func.model.tune.param_range.code .sidebar }
##### Source Code
\

Below is the simplified version of the source code of the [model.tune.param_range](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/common-helper.functions.R#L198) function:
```{r eval=FALSE}
model.tune.param_range <- function(loop_starter,
                             tune_dir_path,
                             cache_file_base_name,
                             fn_tune.test.param_value,
                             # interval_divisor.multiplier = 4,
                             max.identical.min_RMSE.count = 4,
                             endpoint.min_diff = 0, #1e-07,
                             break.if_min = TRUE,
                             steps.beyond_min = 2){

@@
